---
title: "Covid19 vaccine study Fall2021"
author: "Emilee Coronel"
date: "11/26/2021"
output: 
  html_document:
    toc: true
    toc_depth: 5
---


```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
TheData <-read.csv(file="FallFinalData.csv",header=TRUE) %>% select(-"X.1", -"X")
Covid91 <- TheData %>% mutate(across(where(is.character),as.factor)) 
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Part of the information you need to compute is the so-called effectiveness of the drug for each subgroup of people. Include a 95% confidence interval of the effectiveness for each different subgroup. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males

```{r}
Males <- filter(Covid91, sex == "Man")
```


```{r}
barchartGC(~treatment + infected,data=Males)
barchartGC(~treatment + infected,data=Males, type="percent")
```

The bar chart shows us that the ones that who took the placebo got sick more thank the males who took Covid91. We can see that the pink bar for the placebo was almost double in it size since it is compared to the drug. This is indicating that the drug is in effective for the Males group. 

### Numerical Summaries
```{r}
table1 <- xtabs(~infected + treatment,data=Males)
rowPerc(table1)
colPerc(table1)
```

About 33.11% of the sick Males had the treatment, and about 66.89% of the Males that who got sick took the placebo. This is about a 2 to 1 difference since it lets us know that the treatment is in fact working. 

## Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

According to the Chi-squared test, the p-value is very small(2.2e-16). This is indicating that we need to reject the null hypothesis and that there is a difference between the Males who took the placebo and with that Males who took the drug 


```{r}
MaleEstimate <- Males$estimate
Malelow <- Males$conf.int[1]
Malehigh <- Males$conf.int[2]

Maleowprob <- Malelow / (1 + Malelow)
Maleloweffect <- Malelow * 100

Malehigh <- Malehigh / (1 + Malehigh)
Maleloweffect <- Malehigh * 100

Maleprob <- MaleEstimate / (1+ MaleEstimate)
MaleEffect <- Maleprob * 100
options(digits = 1)
```

The effectiveness of the Covid91 vaccine for the Males is 67.2. The confidence that is interval for the males is 63. to 71

# Females

# Drug Users

# LGBTQ

# Conclusions and recommendations

