---
title: "Covid19 vaccine study Fall2021"
author: "Emilee Coronel"
date: "11/26/2021"
output: 
  html_document:
    toc: true
    toc_depth: 5
---


```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
TheData <-read.csv(file="FallFinalData.csv",header=TRUE) %>% select(-"X.1", -"X")
Covid91 <- TheData %>% mutate(across(where(is.character),as.factor)) 
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Part of the information you need to compute is the so-called effectiveness of the drug for each subgroup of people. Include a 95% confidence interval of the effectiveness for each different subgroup. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbiage when referencing p-values in conclusions.

# Males

```{r}
Males <- filter(Covid91, sex == "Man")
```


```{r}
barchartGC(~treatment + infected,data=Males)
barchartGC(~treatment + infected,data=Males, type="percent")
```

The bar chart shows us that the ones that who took the placebo got sick more thank the males who took Covid91. We can see that the pink bar for the placebo was almost double in it size since it is compared to the drug. This is indicating that the drug is in effective for the Males group. 

### Numerical Summaries
```{r}
table1 <- xtabs(~infected + treatment,data=Males)
rowPerc(table1)
colPerc(table1)
```

About 33.11% of the sick Males had the treatment, and about 66.89% of the Males that who got sick took the placebo. This is about a 2 to 1 difference since it lets us know that the treatment is in fact working. 

## Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Malesmod <- fisher.test(table1)
```

According to the Chi-squared test, the p-value is very small(2.2e-16). This is indicating that we need to reject the null hypothesis and that there is a difference between the Males who took the placebo and with that Males who took the drug 


```{r}
MaleEstimate <- Malesmod$estimate
Malelow <- Malesmod$conf.int[1]
Malehigh <- Malesmod$conf.int[2]

Malelowprob <- Malelow / (1 + Malelow)
Maleloweffect <- Malelowprob * 100

Malehighprob <- Malehigh / (1 + Malehigh)
Maleloweffect <- Malehighprob * 100

Maleprob <- MaleEstimate / (1+ MaleEstimate)
MaleEffect <- Maleprob * 100
options(digits = 1)
```
The effectiveness of the Covid91 vaccine for Males is 67.2 The confidence interval for Males is 63.1 to 71

# Females

```{r}
Females <- filter(Covid91, sex == "Woman")
```
```{r}
barchartGC(~treatment + infected,data=Females)
barchartGC(~treatment + infected,data=Females, type="percent")
```

According to the bar chart, this vaccine isn't effective for Females. The blue bar is indicating the out of those who  are sick more than 50% had taken the vaccine. 

### Numerical Summaries 
```{r}
table1 <- xtabs(~infected + treatment,data=Females)
rowPerc(table1)
colPerc(table1)
```
This table is a reassures that the Females who got sick, 58% of them had taken the vaccine, versus 42% of the sick that took the placebo. Based on this information we can see that the vaccine seems not to be effective for the Females.

## Inferential
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
Femalesmod <- fisher.test(table1)
```

Since our p-value is very small, we have to reject the null hypothesis, which is indicating that there is a difference between our data. In this case, it looks like the data is different in a bad way, as in the drug is not as effective compared to the placebo.

## Calculating Effectiveness

```{r}
FemaleEstimate <- Femalesmod$estimate
Femalelow <- Femalesmod$conf.int[1]
Femalehigh <- Femalesmod$conf.int[2]

Femalelowprob <- Femalelow / (1 + Femalelow)
Femaleloweffect <- Femalelowprob * 100

Femalehighprob <- Femalehigh / (1 + Femalehigh)
Femaleloweffect <- Femalehighprob * 100

Femaleprob <- FemaleEstimate / (1+ FemaleEstimate)
FemaleEffect <- Femaleprob * 100
options(digits = 1)
```

The effectiveness of the Covid91 vaccine for Females is 40.4 The confidence interval for Females is 35.1 to 45.9

# Drug Users
```{r}
DrugUsers <- filter(Covid91, DrugUser == "yes")
```
```{r}
barchartGC(~treatment + infected,data=DrugUsers)
barchartGC(~infected + treatment,data=DrugUsers, type = "percent")
```


### Numerical Summaries
```{r}
table1 <- xtabs(~infected + treatment,data=DrugUsers)
rowPerc(table1)
colPerc(table1)
```
## Inferential
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
DrugUsersmod <- fisher.test(table1)
```

## Calculating Effectiveness

```{r}
DrugUsersEstimate <- DrugUsersmod$estimate
Druglow <- DrugUsersmod$conf.int[1]
Drughigh <- DrugUsersmod$conf.int[2]

Druglowprob <- Druglow / (1 + Druglow)
Drugloweffect <- Druglowprob * 100

Drughighprob <- Drughigh / (1 + Drughigh)
Drugloweffect <- Drughighprob * 100

Drugprob <- DrugUsersEstimate / (1+ DrugUsersEstimate)
DrugUsersEffect <- Drugprob * 100
options(digits = 1)
```

The effectiveness of the Covid91 vaccine for drug users is 64.7. The confidence interval for drug users is 55 to 73.6.


# LGBTQ
```{r}
LGBTQ <- filter(Covid91, LGBTQ == "gay")
```
```{r}
barchartGC(~treatment + infected,data=LGBTQ)
barchartGC(~treatment + infected,data=LGBTQ, type="percent")
```
According to the bar chart, the LGBTQ does not benefit from taking this drug. In our sick group, more than 50% had taken the drug. 

### Numerical Summaries 

```{r}
table1 <- xtabs(~infected + treatment,data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```

This chart tells us that the LGBTQ should not take this drug, as 71% of those who were sick had taken the drug versus 29% of the sick people had not. 

## Inferential
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
LGBTQmod <- fisher.test(table1)
```

According to the chi-squared test, the p-value is very small (6e-06). This indicated that we need to reject the null hypothesis and that there is a difference between the LGBTQ who took the placebo and LGBTQ who took the drug. In this case the difference is in an ineffective way, as there are more people who got infected while on the drug. 


## Calculating Effectiveness

```{r}
LGBTQEstimate <- LGBTQmod$estimate
LGBTQlow <- LGBTQmod$conf.int[1]
LGBTQhigh <- LGBTQmod$conf.int[2]

LGBTQlowprob <- LGBTQlow / (1 + LGBTQlow)
LGBTQloweffect <- LGBTQlowprob * 100

LGBTQhighprob <- LGBTQhigh / (1 + LGBTQhigh)
LGBTQloweffect <- LGBTQhighprob * 100

LGBTQprob <- LGBTQEstimate / (1+ LGBTQEstimate)
LGBTQEffect <- LGBTQprob * 100
options(digits = 1)
```

The effectiveness of Covid91 vaccine for LGBTQ is 28.6. The confidence interval for those in the LGBTQ community is 20.5 to 38.

## Cousulsion and recommendations
